Crunching Large Graphs with Commodity Processors

### 2011-04-20 doc/hotpar2011/softxmt-hotpar2011.tex
fix spacings
grammer intro

Many important emerging applications such as social network analysis,
bioinformatics, and sensor networks rely on crunching very large
graphs. Unfortunately, the computational cost of these applications
gets quickly out of hand. Tools to analyze social networks and query
semantic graphs with billions of entities exist
today~\cite{joslyn:hpsf, ediger:msda, brocheler:cosi, pegasus,
  pregel}, but graphs of interest to defense applications are expected
have trillions of entities~\cite{hartley:mssg, kolda:imkd}. Speeding
up these applications at a low cost would have a large impact in how
we deal with large data-sets and make them even more valuable.

@@

This paper describes the vision of our system and explores the
feasibility of our ideas using a single-node implementation of a
lightweight threading runtime. We use prefetch instructions together
with lightweight threads to tolerate the latency of DRAM access on the
local node. We use worst-case pointer chasing benchmarks to
verify our runtime's overhead is acceptable. We also model the effects
of pointer chasing on a remote node by artificially increasing
latencies to several microseconds.

@@

\begin{figure*}[ht]
\vspace{-0.2in}
%list walk ILP
\begin{minipage}[b]{0.3\linewidth}{\small
\centering
\begin{alltt}{\scriptsize
  while (count-- > 0) \{
    list1 = list1->next;
    list2 = list2->next;
    \ldots
    list\(n\) = list\(n\)->next;
  \}
  }\centering{\bf (a)}
\end{alltt}
%\caption{Pseudocode for pointer chasing without coroutines.}
\label{fig:pointernocoro}
}\end{minipage}
%listwalk green threads
\begin{minipage}[b]{0.35\linewidth}{\small
\centering
\begin{alltt}{\scriptsize
  while (count-- > 0) \{
     readIssue(&(list->next));
     yield();
     list = readCompete(&(list->next));
 \}
 }\centering{\bf (b)}
\end{alltt}
  %%n lists
  % while (count-- > 0) \{
  % prefetch(&(list1->next));
  %   prefetch(&(list2->next));
   %  \ldots
   %  prefetch(&(list\(n\)->next));
   %  switch();
   %  list1 = read(&(list1->next));
   %  list2 = read(&(list2->next));
   %  \ldots
   %  list\(n\) = read(&(list\(n\)->next));
%\caption{Pseudocode for pointer chasing using coroutines.}
\label{fig:pointercoro}
}\end{minipage}
%cache pressure
\begin{minipage}[b]{0.32\linewidth}{\small
\centering
\begin{alltt}{\scriptsize
  while (count-- > 0) \{
    readIssue(&(remoteList->next));
    yield();
    remoteList = readComplete(&(remoteList->next));
    for( i in 1 to num_local_updates ) \{
      localList->data++;
      localList = localList->next;
   \} \}
   }\centering{\bf (c)}
\end{alltt}
%\caption{Pseudocode for pointer chasing with local updates.}
\label{fig:pointerupdate}
}\end{minipage}
%\vspace{5pt}
\caption{Pseudocode for: (a) pointer chasing without coroutines, (b) pointer chasing using coroutines, (c) pointer chasing with local updates. Coroutine code, as in (b), might also include ILP like what (a) depends on for memory concurrency, e.g., by doing multiple prefetches to independent lists, switch, then multiple reads.}
\label{fig:code}
\end{figure*}

@@

% cache pressure figure
\begin{figure}[h]
  \begin{center}
    \includegraphics[width=0.45\textwidth]{figures/cachepressurenewview-edited.pdf}
  \end{center}
  \vspace{0.0in}
  \caption{Cache pressure with coroutines, on six cores. The vertical axis is remote reference rate. (a) with 8 coroutines per core. (b) with 48 coroutines per core, about the number required for the network rate in simulated delay experiments. }
  \label{fig:cache-pressure}
\end{figure}

>>>>

Many important emerging applications such as social network analysis,
bioinformatics, and sensor networks rely on crunching very large
graphs. Unfortunately, the computational cost of these applications
gets quickly out of hand. Tools to analyze social networks and query
semantic graphs with billions of entities exist
today~\cite{joslyn:hpsf, ediger:msda, brocheler:cosi, pegasus,
  pregel}, but graphs of interest to defense applications are expected
to have trillions of entities~\cite{hartley:mssg, kolda:imkd}. Speeding
up these applications at a low cost would have a large impact in how
we deal with large data-sets and make them even more valuable.

@@

This paper describes the vision of our system and explores the
feasibility of our ideas using a single-node implementation of a
lightweight threading runtime. We use prefetch instructions together
with lightweight threads to tolerate the latency of DRAM access on the
local node. We use worst-case pointer chasing benchmarks to
verify our runtime's overhead is acceptable. We also model the effects
of pointer chasing with a remote node by artificially increasing
latencies to several microseconds.

@@

\begin{figure*}[ht]
\vspace{-0.2in}
%list walk ILP
\begin{minipage}[b]{0.3\linewidth}{\small
\centering
\begin{alltt}{\scriptsize
  while (count-- > 0) \{
    list1 = list1->next;
    list2 = list2->next;
    \ldots
    list\(n\) = list\(n\)->next;
  \}
  }\centering{\bf (a)}
\end{alltt}
%\caption{Pseudocode for pointer chasing without coroutines.}
\label{fig:pointernocoro}
}\end{minipage}
%listwalk green threads
\begin{minipage}[b]{0.35\linewidth}{\small
\centering
\begin{alltt}{\scriptsize
  while (count-- > 0) \{
     readIssue(&(list->next));
     yield();
     list = readCompete(&(list->next));
 \}
 }\centering{\bf (b)}
\end{alltt}
  %%n lists
  % while (count-- > 0) \{
  % prefetch(&(list1->next));
  %   prefetch(&(list2->next));
   %  \ldots
   %  prefetch(&(list\(n\)->next));
   %  switch();
   %  list1 = read(&(list1->next));
   %  list2 = read(&(list2->next));
   %  \ldots
   %  list\(n\) = read(&(list\(n\)->next));
%\caption{Pseudocode for pointer chasing using coroutines.}
\label{fig:pointercoro}
}\end{minipage}
%cache pressure
\begin{minipage}[b]{0.32\linewidth}{\small
\centering
\begin{alltt}{\scriptsize
  while (count-- > 0) \{
    readIssue(&(remoteList->next));
    yield();
    remoteList = readComplete(&(remoteList->next));
    for( i in 1 to num_local_updates ) \{
      localList->data++;
      localList = localList->next;
   \} \}
   }\centering{\bf (c)}
\end{alltt}
%\caption{Pseudocode for pointer chasing with local updates.}
\label{fig:pointerupdate}
}\end{minipage}
\vspace{5pt}
\caption{Pseudocode for: (a) pointer chasing without coroutines, (b) pointer chasing using coroutines, (c) pointer chasing with local updates. Coroutine code, as in (b), might also include ILP like what (a) depends on for memory concurrency, e.g., by doing multiple prefetches to independent lists, switch, then multiple reads.}
\label{fig:code}
\end{figure*}

@@

% cache pressure figure
\begin{figure}[h]
  \begin{center}
    \includegraphics[width=0.49\textwidth]{figures/cachepressurenewview-edited.pdf}
  \end{center}
  \vspace{0.0in}
  \caption{Cache pressure with coroutines, on six cores. The vertical axis is remote reference rate. (a) with 8 coroutines per core. (b) with 48 coroutines per core, about the number required for the network rate in simulated delay experiments. }
  \label{fig:cache-pressure}
\end{figure}
